---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mori <a href="https://mori.r-lib.org"><img src="man/figures/logo.svg" align="right" height="139" alt="mori website" /></a>

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/mori)](https://CRAN.R-project.org/package=mori)
[![R-CMD-check](https://github.com/r-lib/mori/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/r-lib/mori/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/r-lib/mori/graph/badge.svg)](https://app.codecov.io/gh/r-lib/mori)
<!-- badges: end -->

[![Ask DeepWiki](https://deepwiki.com/badge.svg)](https://deepwiki.com/r-lib/mori)

Shared Memory for R Objects

→ `share()` writes an R object into shared memory and returns a shared version

→ Compact ALTREP serialization — shared objects move through `serialize()` and `mirai()` as small references

→ Lazy access and automatic cleanup — data is read on demand and freed by R's garbage collector

→ OS-level shared memory (POSIX / Win32) — pure C, no external dependencies

<br />

## Installation

``` r
install.packages("mori")
```

## Why mori

<a href="#why-mori"><img src="man/figures/mori-diagram.svg" alt="Diagram showing share() writing an object once into OS-backed shared memory, which is then memory-mapped by other processes using zero-copy ALTREP wrappers" width="720" /></a>

Parallel computing multiplies memory.
When 8 workers each need the same 200 MB dataset, that is 1.6 GB of serialization, transfer, and deserialization.
RAM holds 8 separate copies.

`share()` writes the data into shared memory once.
Each worker then maps the same physical pages.
Per-worker copies become per-worker references.

```{r}
#| label: bench-setup
library(mori)
library(mirai)
library(lobstr)

daemons(8)

# 200 MB data frame — 5 columns × 5M rows
df <- as.data.frame(matrix(rnorm(25e6), ncol = 5))
shared_df <- share(df)
```

Without mori, each worker holds the full data frame.
With mori, each worker holds a small reference into the shared memory region:

```{r}
#| label: bench-mem
mirai_map(1:8, \(i, data) format(lobstr::obj_size(data)),
          .args = list(data = df))[.flat] |> unique()

mirai_map(1:8, \(i, data) format(lobstr::obj_size(data)),
          .args = list(data = shared_df))[.flat] |> unique()
```

The workers also skip 8 × 200 MB of serialization and deserialization, which gives a significant runtime saving:

```{r}
#| label: bench-time
boot_mean <- \(i, data) colMeans(data[sample(nrow(data), replace = TRUE), ])

# Without mori — each daemon deserializes a full copy
mirai_map(1:8, boot_mean, .args = list(data = df))[] |> system.time()

# With mori — each daemon maps the same shared memory
mirai_map(1:8, boot_mean, .args = list(data = shared_df))[] |> system.time()

daemons(0)
```

## Usage

Workers must run on the same machine, because mori shares physical RAM.

### Sharing by name

`shared_name()` returns the shared memory name of a shared object.
`map_shared()` opens a region by this name.
This passes a reference between processes without serialization:

```{r}
#| label: by-name
x <- share(rnorm(1e6))

shared_name(x)
```
```{r}
# Another process can map the region by name
y <- map_shared(shared_name(x))
identical(x[], y[])
```

### Sharing through serialization

The ALTREP serialization hooks emit the same identifier on the wire.
The serialized form is a few bytes, regardless of the data size:

```{r}
#| label: serialize
length(serialize(x, NULL))
```

This is transparent to any R serialization pathway.
`mirai`, `parallel`, `callr`, and base R `serialize()` all carry shared objects as references, not copies.

Sub-elements of a shared list serialize as references too.
In this case, each element travels as a path into the parent shared region, not as the full data:

```{r}
#| label: mirai-map
daemons(3)

# Share a list — all 3 vectors in a single shared region
lst <- share(list(a = rnorm(1e6), b = rnorm(1e6), c = rnorm(1e6)))

# Each element arrives on the worker as a zero-copy reference
mirai_map(lst, \(v) format(lobstr::obj_size(v)))[.flat] |> unique()

daemons(0)
```

## How It Works

### What gets shared

`share()` writes all atomic vector types, lists, and data frames directly into shared memory.
Attributes are preserved end-to-end.
Pairlists become lists.
The returned ALTREP wrappers point into the shared memory region.
There is no deserialization and no per-process memory allocation.

`share()` returns all other R objects (environments, closures, language objects) unchanged.
It creates no shared memory region for them.

### Lazy access

A data frame uses a single shared region.
Workers read columns on demand.
A worker that needs 3 of 100 columns loads only 3.
Character strings load on demand, one element at a time.

```{r}
#| label: lazy
df <- share(as.data.frame(matrix(rnorm(1e7), ncol = 100)))
shared_name(df)        # one region for all 100 columns
shared_name(df[[50]])  # sub-path into the same region
```

### Lifetime

R's garbage collector manages the shared memory.
A region stays alive while R holds a reference to any object backed by it.
The reference can be the original from `share()`, or a column or sub-list extracted from it.
This reference can be in the original or another process.
When no references remain, or when the session exits cleanly, R frees the shared memory automatically.

**Important:** Make sure that R does not garbage-collect the return value of `share()` before a consumer maps its shared memory.

If a process dies before cleanup runs (a crash, `SIGKILL`, or the OOM killer), its region can be left behind.
`prune_shared()` reclaims these orphans.
It removes only regions whose creating process is no longer running.

### Copy-on-write

Consumers map the shared data read-only.
This prevents corruption of the shared region.
Changes are always local.
Copy-on-write makes sure that other processes continue to read the original shared data:

- **Structural changes** to a list or data frame (add, remove, or reorder elements) produce a regular R list.
  The shared region stays unchanged.
- **Modifying values** of a shared vector (for example, `X[1] <- 0`) materializes only that vector into a private copy.
  Other vectors in the same shared region stay zero-copy.

--

The mori project is released with a [Contributor Code of Conduct](https://mori.r-lib.org/CODE_OF_CONDUCT.html).
When you contribute to this project, you agree to obey its terms.
